---
title: "scLean API Reference"
author: "rzyang-dev"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{API Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## Key Concepts

| Concept | Description |
|---------|-------------|
| **Disk-backed layers** | Assay data (counts, data, scale.data) stored in HDF5 files, not RAM |
| **Chunked computation** | All operators read/write in adaptive chunks controlled by the resource scheduler |
| **Auto chunk sizing** | Chunk sizes adapt automatically to available RAM; override with `SetChunkSize()` |
| **OOM degradation** | 3-level recovery: dense path → sparse path → shrink chunk → retry |
| **Seurat compatibility** | API identical to Seurat — S3 dispatch routes to disk-backed methods |

---

## Data I/O & Conversion

```{r}
fn <- c(
  "`LoadScleanObject(input, hdf5_path, assay, ...)`",
  "`as.scLean(object, hdf5_path, assay, layers, ...)`",
  "`as.Seurat.scLean(object, layers, assay, ...)`",
  "`CreateSCleanAssay(counts, hdf5_path, assay, ...)`",
  "`CreateSCleanAssayFromHDF5(hdf5_path, assay)`",
  "`HDF5BackedMatrix$new(hdf5_path, group_path)`"
)
desc <- c(
  "Load from 10X dir, dgCMatrix, HDF5, or Seurat object",
  "Convert Seurat object to disk-backed storage",
  "Convert disk-backed layers back to in-memory Seurat",
  "Create disk-backed assay from a dgCMatrix",
  "Reopen assay from an existing HDF5 file",
  "R6 matrix proxy backed by HDF5; dim, [, t, %*%"
)
help <- c(
  "`help(LoadScleanObject)`",
  "`help(as.scLean)`",
  "`help(as.Seurat.scLean)`",
  "`help(CreateSCleanAssay)`",
  "`help(CreateSCleanAssayFromHDF5)`",
  "`help(HDF5BackedMatrix)`"
)

io_df <- data.frame(Function = fn, Description = desc, Help = help)
knitr::kable(io_df, caption = "Data I/O & Conversion")
```

---

## Analysis Pipeline

```{r}
fn_pipe <- c(
  "`NormalizeData(object, method, scale.factor, ...)`",
  "`ScaleData(object, features, do.scale, do.center, ...)`",
  "`FindVariableFeatures(object, nfeatures, ...)`",
  "`RunPCA(object, npcs, reduction.name, ...)`",
  "`FindNeighbors(object, reduction, dims, k.param, ...)`",
  "`FindClusters(object, algorithm, resolution, n.iter, ...)`",
  "`FindMarkers(object, ident.1, ident.2, test.use, ...)`",
  "`FindAllMarkers(object, test.use, logfc.threshold, min.pct, ...)`"
)
desc_pipe <- c(
  "Chunked normalization: LogNormalize, CLR, or RelativeCounts",
  "Per-gene mean/sd (Welford); PCA applies centering on-the-fly",
  "VST-based variable gene selection with sparse fallback",
  "IRLBA PCA from HDF5 data; on-the-fly centering",
  "KNN/SNN graph; Annoy for large N, brute-force when memory-bound",
  "Leiden or Louvain clustering on SNN graph",
  "Differential expression: Wilcoxon, t-test, or logistic regression",
  "Per-cluster DE; wraps FindMarkers for all clusters"
)
help_pipe <- c(
  "`help(NormalizeData.scLeanAssay)`",
  "`help(ScaleData.scLeanAssay)`",
  "`help(FindVariableFeatures.scLeanAssay)`",
  "`help(RunPCA.Seurat)`",
  "`help(FindNeighbors.Seurat)`",
  "`help(FindClusters.Seurat)`",
  "`help(FindMarkers.Seurat)`",
  "`help(FindAllMarkers.scLeanAssay)`"
)

pipe_df <- data.frame(Function = fn_pipe, Description = desc_pipe, Help = help_pipe)
knitr::kable(pipe_df, caption = "Analysis Pipeline")
```

---

## Integration

```{r}
fn_int <- "`IntegrateData.scLean(object, batch, n.ccs, n.mnn, sigma, max.iter, ...)`"
desc_int <- "MNN-based batch correction with Gaussian kernel smoothing"
help_int <- "`help(IntegrateData.scLean)`"

int_df <- data.frame(Function = fn_int, Description = desc_int, Help = help_int)
knitr::kable(int_df, caption = "Integration")
```

---

## Resource Control

```{r}
fn_ctl <- c(
  "`SetChunkSize(object, size)`",
  "`SetMaxRAM(max_ram = 2048)`",
  "`SetThreads(n = NULL)`",
  "`SetVerbose(verbose = TRUE)`",
  "`MemoryReport(object = NULL)`",
  "`CheckResources()`",
  "`RefreshMemory(object = NULL)`",
  "`PerformanceSnapshot()`"
)
desc_ctl <- c(
  "Set cells-per-chunk; NA to restore auto",
  "Set RAM limit in MB (default: 2048)",
  "Set OpenMP thread count; NULL to restore single-threaded",
  "Toggle progress messages",
  "Full resource report (RAM, CPU, pressure, HDF5 info)",
  "Warn if memory or CPU pressure > 90%",
  "Trigger R garbage collection",
  "Quick RSS + wall-time snapshot for timing"
)
help_ctl <- c(
  "`help(SetChunkSize)`",
  "`help(SetMaxRAM)`",
  "`help(SetThreads)`",
  "`help(SetVerbose)`",
  "`help(MemoryReport)`",
  "`help(CheckResources)`",
  "`help(RefreshMemory)`",
  "`help(PerformanceSnapshot)`"
)

ctl_df <- data.frame(Function = fn_ctl, Description = desc_ctl, Help = help_ctl)
knitr::kable(ctl_df, caption = "Resource Control & Monitoring")
```

---

## See Also

- `vignette("sclean", package = "scLean")` — step-by-step tutorial
- `help(package = "scLean")` — complete function reference
