---
title: "Getting Started with scLean"
author: "rzyang-dev"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with scLean}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r load}
library(scLean)
library(Seurat)

# Create synthetic test data
set.seed(42)
counts <- Matrix::rsparsematrix(500, 300, density = 0.15,
  dimnames = list(paste0("gene", 1:500), paste0("cell", 1:300)))
obj <- LoadScleanObject(counts, assay = "RNA")
```

## Overview

scLean is a memory-efficient single-cell RNA-seq analysis package that replaces
Seurat's in-memory computation with HDF5-backed, out-of-core algorithms. It is
designed for large datasets (500K+ cells) on machines with limited RAM (8 GB).

```{r overview}
obj
```

## Loading Data

### From a dgCMatrix

```{r load-matrix}
counts2 <- Matrix::rsparsematrix(200, 100, density = 0.1,
  dimnames = list(paste0("g", 1:200), paste0("c", 1:100)))
obj2 <- LoadScleanObject(counts2, assay = "RNA")
obj2
```

### From an existing Seurat object

```{r load-seurat, eval=FALSE}
seurat_obj <- SeuratObject::CreateSeuratObject(counts)
obj <- as.scLean(seurat_obj, hdf5_path = "my_data.h5")
```

### From a 10X Genomics directory

```{r load-10x, eval=FALSE}
pbmc <- LoadScleanObject("path/to/filtered_feature_bc_matrix")
```

## Standard Seurat Pipeline

All functions work the same way as standard Seurat, but computation is
disk-backed and uses adaptive chunk sizing.

### Normalize

```{r normalize}
t0 <- PerformanceSnapshot()
obj <- NormalizeData(obj, normalization.method = "LogNormalize")
t1 <- PerformanceSnapshot()
cat(sprintf("NormalizeData: %.2f seconds\n",
  (t1$wall_time_ns - t0$wall_time_ns) / 1e9))
```

### Find Variable Features

```{r vst}
t0 <- PerformanceSnapshot()
obj <- FindVariableFeatures(obj, nfeatures = 100)
t1 <- PerformanceSnapshot()
cat(sprintf("FindVariableFeatures: %.2f seconds, %d features selected\n",
  (t1$wall_time_ns - t0$wall_time_ns) / 1e9,
  length(VariableFeatures(obj))))
head(VariableFeatures(obj))
```

### Scale

```{r scale}
t0 <- PerformanceSnapshot()
obj <- ScaleData(obj)
t1 <- PerformanceSnapshot()
cat(sprintf("ScaleData: %.2f seconds\n",
  (t1$wall_time_ns - t0$wall_time_ns) / 1e9))
```

### Run PCA

```{r pca}
t0 <- PerformanceSnapshot()
obj <- RunPCA(obj, npcs = 20)
t1 <- PerformanceSnapshot()
cat(sprintf("RunPCA: %.2f seconds\n",
  (t1$wall_time_ns - t0$wall_time_ns) / 1e9))
```

### Find Neighbors and Clusters

```{r neighbors}
t0 <- PerformanceSnapshot()
obj <- FindNeighbors(obj, dims = 1:20)
obj <- FindClusters(obj, resolution = 0.8, algorithm = "leiden")
t1 <- PerformanceSnapshot()
cat(sprintf("FindNeighbors + FindClusters: %.2f seconds\n",
  (t1$wall_time_ns - t0$wall_time_ns) / 1e9))
table(Idents(obj))
```

### Run UMAP

```r
# UMAP delegates to Seurat's implementation (data read from HDF5)
obj <- RunUMAP(obj, dims = 1:20)
DimPlot(obj, reduction = "umap", label = TRUE)
```

## Finding Markers

```{r markers}
t0 <- PerformanceSnapshot()
markers <- FindMarkers(obj, ident.1 = "1", ident.2 = "2")
t1 <- PerformanceSnapshot()
cat(sprintf("FindMarkers: %.2f seconds, %d significant genes\n",
  (t1$wall_time_ns - t0$wall_time_ns) / 1e9,
  nrow(markers)))
head(markers[order(markers$p_val_adj), ], n = 6)
```

```{r all-markers}
all_markers <- FindAllMarkers(obj)
names(all_markers)
```

## Integration

MNN-based batch correction for removing technical variation between samples:

```{r integration}
# Add synthetic batch labels
obj$batch <- sample(c("A", "B"), ncol(obj), replace = TRUE)

t0 <- PerformanceSnapshot()
obj <- IntegrateData.scLean(obj, batch = "batch", n.ccs = 15)
t1 <- PerformanceSnapshot()
cat(sprintf("IntegrateData: %.2f seconds\n",
  (t1$wall_time_ns - t0$wall_time_ns) / 1e9))
```

## Performance Monitoring

scLean includes built-in resource monitoring. Use these between pipeline steps
to detect bottlenecks early.

```{r monitoring}
# System pressure check (warns if memory or CPU > 90%)
CheckResources()

# Detailed resource report
MemoryReport(obj)
```

## Memory Management

```{r memory}
# Check current memory usage
report <- MemoryReport(obj)
cat(sprintf("Free RAM: %.0f MB\n", report$free_ram_mb))
cat(sprintf("Current RSS: %.0f MB\n", report$current_rss_mb))

# Limit to 2 GB
SetMaxRAM(2048)

# Set fixed chunk size
SetChunkSize(obj, 500)

# Restore automatic mode
SetChunkSize(obj, NA)

# Force garbage collection
RefreshMemory(obj)

# Control parallelism
SetThreads(2)
SetThreads()  # restore default
```

## Converting Back

```{r convert}
# Load data layer into memory
obj_mem <- as.Seurat.scLean(obj, layers = "data")
obj_mem
```

## Saving and Loading

scLean objects store only metadata and the HDF5 path in memory. The HDF5 file
must exist at the same path when reloading:

```{r save-load, eval=FALSE}
saveRDS(obj, file = "sclean_obj.rds")
obj <- readRDS("sclean_obj.rds")
```

## See Also

- `vignette("reference", package = "scLean")` — API cheat sheet
- `help(package = "scLean")` — complete function reference

## Session Info

```{r}
sessionInfo()
```
